E1: ENRON Sentiment

### 2016-03-25 paper/paper.tex
done project setup

@@

\section{Project setup}
\label{sec:p-s}
TODO

>>>>

\usepackage{hyperref}
\hypersetup{pdftex,colorlinks=true,allcolors=blue}
\usepackage{hypcap}

@@

\section{Project setup}
\label{sec:p-s}
In this section we explain our software setup and the developed architecture.
We created and deployed three Spark jobs, each one with a specific task, in
order to render them more easily maintainable, manageable and deployable.

\subsection{Archive Extraction}
\label{sub-sec:archive-extraction}
The first step in our data analysis pipeline comprises the extraction of the
dataset from its \textit{zipped} form. The dataset consists of a collection of
archives, each one enclosing an Enron's employee mailbox, containing emails in
both native (eml, pdf, docx) and plain text (in which attachments have been
converted into plain text too). We decided to extract and work with the plain
text messages, filtering out attachments. The \texttt{UnzipperDriver} job is
responsible for this. The goal of the job is to unzip the archives and store a
collection of Scala tuples of the kind \texttt{(mailboxName, documents)} where
the first element is the mailbox name (extracted from the zip filename) and
documents is a \texttt{Seq} of extracted documents.

\subsection{ETL}
\label{sub-sec:etl}
Second step is the \textit{ETL} (Extract, Transform, Load). Here the term
\textit{Extraction} has a different meaning than that of the previous job.
The extraction part in \texttt{ETLDriver} is about extracting data from emails
in a structured form, that is we need to extract emails body and headers from
the raw data. Do do this we created an \texttt{EmailParser} object that
implements the extraction and parsing logic. In this we apply a series of
transformations to the raw emails:
\begin{itemize}
	\item extract email headers (\textit{Date}, \textit{From}, \textit{To},
		\textit{Cc}, \textit{Bcc} and \textit{Subject});
	\item separate body from headers;
	\item clean body from common dataset footer \footnote{Every email contains a
		disclaimer from EDRM, the company that cleansed and published this
		version of the Enron Corpus};
	\item clean body from quoted emails, ``\textit{Original Message}'' and
		``\textit{Forwarded By}'' text and delimiters.
\end{itemize}

Because we are interested only in emails exchanged by Enron employees, we need a
way to retrieve and identify those people from the email headers. Unfortunately
this is not straightforward because addresses and people names have different
formats like ``\textit{first-name last-name}'', ``\textit{last-name, first-name}'',
``\textit{first-name last-name, email}'' or ``\textit{last-name, initial}''.
To solve this problem we use an external list of \textit{mailbox custodians}
that gives us first, last name and its role within Enron. We then search
\textit{From}, \textit{To}, \textit{Cc} and \textit{Bcc} headers for known
custodians. If for an email it fails to find any known custodian, the message is
discarded.

Next processing step consists in transforming the extracted dataset into a form
more easily manageable and queryable. We use CoreNLP to tokenize and find
sentiment at sentence level in emails. Because the sentiment is at sentence
level we need a way to aggregate those into a single sentiment for the whole
email. To aggregate sentiments we use the following formula:
\begin{equation}
	S = \frac{\sum_{s \in SS} s}{|SS|}
	\label{eq:sentiment-aggregation}
\end{equation}
where $SS$ is the set of sentence sentiments.

Last step consists in storing the dataset for future analysis. We split the
dataset in two: one contains the full email body and no sentiment, the other
does not contain the body but has sentiments. We do this in order to have a more
lightweight, sentiment-tagged corpus that we can analyse with more agility. We
convert both datasets into Spark SQL's \texttt{DataFrame} that enables us to
perform queries and data aggregation using the SQL language.

\subsection{Sentiment Analyser}
\label{sub-sec:sentiment-analyser}
Last job that we use in our pipeline is the \texttt{SentimentResumer}. It is
responsible for querying the sentiment-tagged dataset and joins it with Enron
stock prices based on date. To do this we group emails by date and then we
average the sentiments for that day and then join them with the stock prices
data on the date field. Lastly we store the results as JSON so that we can
then download them and use them for visualization purposes.

\subsection{Visualization}
\label{sub-sec:visualization}
The resulting JSON file is used to visualize the correlation of Enron stock
prices with email sentiment. We created a simple line chart with two lines and
two vertical axes. A line represents the stock prices and another represents
email sentiments. We built the interface in HTML and Coffeescript, using the d3
visualization library. Because the codebase is hosted on Github we have chosen
to deploy the visualization using Github Pages
\footnote{\url{http://acidghost.github.io/ENRON-sentiment-analysis/visualization/}}
which offers free hosting of static files.
